Stack Safety as a Security Property

### 2020-07-10 main.tex
CP

We now present our model of enforcement, which we introduced
informally in \Cref{sec:threat}.  A policy is a general model of an
enforcement mechanism, consisting of a set of policy states \(\pol \in
\POLS\), a policy step function \((\mach, \pol) \polstep \pol' :
\MACHS \times \POLS \rightharpoonup \POLS\). We assume that the
compiler provides an initial policy state $\pol_0$ to accompany 
the initial machine state.
%
A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
\Cref{sec:enforcement}.

@@

In this section we formalize our definition of stack safety. We begin
by describing our model of knowledge about the program structure, and
then dive into the formal definitions of local state encapsulation and
well-bracketed control flow that together constitute stack safety.

@@

We say that a system enjoys {\em stack integrity} if for every initial
state $(\mach_0,\pol_0)$ and its corresponding induced trace $\MPT$ (with
$\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT$), for every
$n$-transition from $\mpstate{\mach_c}{\pol_c}$ to
$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, if we take the prefix of the
call until its corresponding return $\mpstate{\mach_e}{\pol_e}
\hookrightarrow \MPT_{call} ~ | ~ \ret{\mach_c}$, then $\MPT_{call}$ must
satisfy trace integrity:
\[\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\]

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
For every initial state \((\mach_0,\pol_0)\), we take the induced trace
\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\). Then for
any \(n\)-transition from $\mpstate{\mach_c}{\pol_c}$ to
$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, we take the prefix of the trace
from the entry point until its corresponding return:
%
\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
%
We also take the corresponding trace from any variant \(\mach_e'
\approx_{\mathit{Cof}(\mach_c,n)} \mach_e\):
%
\[\mach_e' \hookrightarrow \machT_{\mathit{call}} | \ret{\mach_c}\]

@@

If trace confidentiality holds for all such traces,
\(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
\machT_{\mathit{call}}\) and it also holds for the toplevel execution
trace \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
system as a whole enjoys {\em stack confidentiality} with respect to
\(\callmap\).

@@

\definition%{Control Separation}
A system enjoys {\em control separation}
% with respect to a function map
%\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
if, for any initial state \((\mach_0, \pol_0)\) and any adjacent pair of states
\(\mpostate{\mach_1}{\pol_1}{\obs_1}\) and \(\mpostate{\mach_2}{\pol_2}{\obs_2}\)
in the induced trace from \(\mpstate{\mach_0}{\pol_0}\),
if \(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname))\),
then \(\mpstate{\mach_1}{\pol_1}\) is either a call \(\callmap(\mach_1(\PCname))\)
or a return \(\retmap(\mach_1(\PCname))\).

@@

\definition%{Entry Integrity}
A system enjoys {\em entry integrity}
%with respect to a call map
%\(\callmap\) and entry map \(\entmap\)
if, for any initial state \((\mach_0, \pol_0)\)
and any n-transition from \((\mach_c,\pol_c)\) to
\((\mach_e, \pol_e)\) in the induced trace from
\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a valid entry:
\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]

@@

\definition%{Return Integrity}
A system enjoys {\em return integrity} if all of its returns match prior calls
and return to the appropriate location. For any initial state \((\mach_0\, \pol_0)\), let
\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pol_0}\).
\(\MPT\) may not have an unmatched return -- \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
may not hold for any \(\MPT_{tail}\). For any state pair \((\mach_c, \pol_c)\)
corresponding to a call \(\MPT\), if
\(\mpstatename = \mpostate{\mach_r}{\pol_r}{\obs_r}\)
is the first unmatched return such that
\( \Unmatched ~ \MPT ~ \MPT_{tail}\) and \(\head{\MPT_{tail}} = \mpstatename\),
then $\mach_r$ must correspond to a return state for $\mach_c$:
\(\ret{\mach_c}\ \mach_r\)

@@

\definition
%
A system satisfies {\em lockstep stack safety} if for every initial machine $(\mach,\pol)$,
the trace induced from $(\mach, \pol)$ satisfies trace lockstep stack safety with
respect to the empty stack.

@@

      {\em Observable stack confidentiality} is structured like stack
      confidentiality. Let \(\callmap\) be a callmap; for each initial
      state \((\mach_0, \pol_0)\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
      to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
      \(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\), take any
      variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} |
      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} |
      \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
      \ret{\mach_c} ~ \MPT ~ \machT\).

>>>>

We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
A policy is a general model of an enforcement mechanism, consisting of a
set of policy states \(\pol \in \POLS\), a policy step function
\((\mach, \pol) \polstep \pol' : \MACHS \times \POLS
\rightharpoonup \POLS\), and an initializer \(\pinit : \MACHS \rightharpoonup
\POLS\). The initializer establishes the initial policy state for a given
machine state, but is partial to reflect that some policies are only defined
on states that are well-formed in some way.
%
\apt{Is this our current thinking?}
\leo{I thought this was what we settled on in the last two days'
  discussions}\bcp{That was not my impression!  I don't see how the policy
  state can be recovered from just the machine state---e.g., the machine
  state doesn't say where the calls and returns are, does it?!}
%
A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
\Cref{sec:enforcement}.

@@

We are finally ready for our definition of stack safety. We begin
by describing how we model knowledge about the program structure,
then dive into the formal definitions of local state encapsulation and
well-bracketed control flow, which together constitute stack safety.

@@

We say that a system enjoys {\em stack integrity} if for every initial
state $\mach_0$ and its corresponding induced trace $\MPT$ (with
$\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT$), for every
$n$-transition from $\mpstate{\mach_c}{\pol_c}$ to
$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, if we take the prefix of the
call until its corresponding return $\mpstate{\mach_e}{\pol_e}
\hookrightarrow \MPT_{call} ~ | ~ \ret{\mach_c}$, then $\MPT_{call}$ must
satisfy trace integrity:
\[\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\]

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
For every initial state \(\mach_0\), we take the induced trace
\(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\). Then for
any \(n\)-transition from $\mpstate{\mach_c}{\pol_c}$ to
$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, we take the prefix of the trace
from the entry point until its corresponding return:
%
\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
%
We also take the corresponding trace from any variant \(\mach_e'
\approx_{\mathit{Cof}(\mach_c,n)} \mach_e\):
%
\[\mach_e' \hookrightarrow \machT_{\mathit{call}} | \ret{\mach_c}\]

@@

If trace confidentiality holds for all such traces,
\(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
\machT_{\mathit{call}}\) and
\(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
system as a whole enjoys {\em stack confidentiality} with respect to
\(\callmap\).

@@

\definition%{Control Separation}
A system enjoys {\em control separation}
% with respect to a function map
%\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
if, for any initial state \(\mach_0\) and any adjacent pair of states
\(\mpostate{\mach_1}{\pol_1}{\obs_1}\) and \(\mpostate{\mach_2}{\pol_2}{\obs_2}\)
in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\),
if \(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname))\),
then \(\mpstate{\mach_1}{\pol_1}\) is either a call \(\callmap(\mach_1(\PCname))\)
or a return \(\retmap(\mach_1(\PCname))\).

@@

\definition%{Entry Integrity}
A system enjoys {\em entry integrity}
%with respect to a call map
%\(\callmap\) and entry map \(\entmap\)
if, for any initial state \(\mach_0\)
and any n-transition from \((\mach_c,\pol_c)\) to
\((\mach_e, \pol_e)\) in the induced trace from
\(\mpstate{\mach_0}{\pinit(\mach_0)}\), $\mach_e$ is a valid entry:
\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]

@@

\definition%{Return Integrity}
A system enjoys {\em return integrity} if all of its returns match prior calls
and return to the appropriate location. For any initial state \(\mach_0\), let
\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pinit(\mach_0)}\).
\(\MPT\) may not have an unmatched return -- \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
may not hold for any \(\MPT_{tail}\). For any state pair \((\mach_c, \pol_c)\)
corresponding to a call \(\MPT\), if
\(\mpstatename = \mpostate{\mach_r}{\pol_r}{\obs_r}\)
is the first unmatched return such that
\( \Unmatched ~ \MPT ~ \MPT_{tail}\) and \(\head{\MPT_{tail}} = \mpstatename\),
then $\mach_r$ must correspond to a return state for $\mach_c$:
\(\ret{\mach_c}\ \mach_r\)

@@

\definition
%
A system satisfies {\em lockstep stack safety} if for every initial machine $\mach$,
the trace induced from $(\mach, \pinit(\mach))$ satisfies trace lockstep stack safety with
respect to the empty stack.

@@

      {\em Observable stack confidentiality} is structured like stack
      confidentiality. Let \(\callmap\) be a callmap; for each initial
      state \(\mach_0\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
      to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
      \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\), take any
      variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} |
      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} |
      \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
      \ret{\mach_c} ~ \MPT ~ \machT\).
